null_model

### 2020-06-04 analysis/ts_null_model.Rmd
first commit

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(biodivTS)
```

@@

Simulation functions
```{r}
get_plot_sample <- function(data, filter_plot, n){
  plot_data <- data %>% filter(plot_id == filter_plot)

  rich_levels <- plot_data %>%
    select(year, id) %>%
    distinct() %>%
    count(year) %>%
    pull(n) %>%
    unique()

  species_pool <- plot_data %>%
    pull(id) %>%
    na.omit() %>%
    unique() #get column of unique scientific names

  if (length(species_pool) < max(rich_levels)) browser()

@@

  site_samples <- data.frame()
  for (j in 1:n){
    iter_samples <- data.frame()
    for(i in rich_levels){
    iter_samples <- rbind(iter_samples, get_samp_species(i))
    }
    site_samples <- rbind(site_samples, iter_samples %>% mutate(n = j))
  }
  site_samples %>%
    mutate(plot_id = filter_plot, value = 1) %>%
    pivot_wider(names_from = samp_species, values_from = value)
}

@@

get_trait_matrix <- function(data, species_list, trait_list){ 
  traits <- data %>%
    filter(id %in% species_list) %>%
    dplyr::select(id, all_of(trait_list)) %>%
    distinct() %>%
    arrange(id) %>%
    column_to_rownames(var = "id")
}
```

Get dataframe of samples for each plot and corresponding species and trait matrices
```{r}
load(here::here("data", "biotime_traits.rda"))

biotime_traits <- biotime_traits %>%
  mutate(plot_id = paste0(study_id, plot)) %>%
  filter(!is.na(id))

n = 100

sample_occurrence <- map_dfr(unique(biotime_traits$plot_id), get_plot_sample, data = biotime_traits, n = n) %>%
  replace(is.na(.), 0)

#colnames(biotime_traits)[17:64]

get_plot_FD <- function(data, trait_range, occ_df, filter_plot) {
  sample_species_mat <- occ_df %>%
    filter(plot_id == filter_plot) %>%
    select(-c(richness, plot_id, n)) %>% #want only species x site info
    select_if(!colSums(.) %in% c(0, NA)) %>% #remove species that don't occur in the plot
    select(sort(tidyselect::peek_vars())) #sort columns to be in alphabetical order, required for dbFDI()

  traits <- colnames(data)[trait_range]
  sample_trait_mat <-
    clean_trait_matrix(get_trait_matrix(data = data, colnames(sample_species_mat), traits))

@@

  df <- occ_df %>%
    filter(plot_id == filter_plot) %>%
    select(richness, plot_id, n) %>%
    bind_cols(sample_fd) %>%
    mutate(merge_test = case_when(richness == nbsp ~ TRUE,
                                  TRUE ~ FALSE))

  if (sum(df$merge_test) != dim(df)[1])
    warning("richness and region columns may not have been appropriately joined with FD data")

  return(df %>% select(-merge_test))
}

test <- get_plot_FD(biotime_traits, 17:64, sample_occurrence, 5617)
```

BAT
```{r}
library(BAT)

get_plot_BAT <- function(data, trait_range, occ_df, filter_plot) {
  sample_species_mat <- occ_df %>%
    filter(plot_id == filter_plot) %>%
    select(-c(richness, plot_id, n)) %>% #want only species x site info
    select_if(!colSums(.) %in% c(0, NA)) %>% #remove species that don't occur in the plot
    select(sort(tidyselect::peek_vars())) #sort columns to be in alphabetical order, required for dbFDI()

  traits <- colnames(data)[trait_range]
  sample_trait_mat <-
    clean_trait_matrix(get_trait_matrix(data = data, colnames(sample_species_mat), traits))

  #get distance matrix
  dist_traits <-  gowdis(sample_trait_mat)
  trait.pco <- dudi.pco(quasieuclid(dist_traits), scannf = FALSE, full = TRUE)

}

test <- get_plot_BAT(biotime_traits, 17:64, sample_occurrence, 5617)

ower.mat_spider <- gower.dist(traits)
euc.pco <- pco(gower.mat_spider,k=3)

barplot(euc.pco$eig)
(euc.pco$eig[1]+euc.pco$eig[2]+euc.pco$eig[3])/sum(euc.pco$eig[euc.pco$eig > 0]) #0.955 variance explained by first 3 axes

traits <- euc.pco$points
row.names(traits) <- row.names(traits2)

>>>>

```{r}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

Simulation functions
```{r}
load(here::here("data", "biotime_traits.rda"))

get_plot_sample <- function(data, study, plot_id, n){
  rich_levels <- data %>%
    filter(study_id == study, plot == plot_id) %>%
    select(year, id) %>%
    distinct() %>%
    count(year) %>%
    pull(n) %>%
    unique()

  species_pool <- data %>% 
    filter(study_id == study, plot == plot_id) %>%
    pull(id) %>%
    unique() #get column of unique scientific names

@@

  site_samples <- data.frame()
  for (j in 1:n){
    iter_samples <- data.frame()
    for(i in rich_levels){
    iter_samples <- rbind(iter_samples, get_samp_species(i))
    }
    site_samples <- rbind(site_samples, iter_samples %>% mutate(n = j))
  }
  site_samples %>%
    mutate(study_id = study, plot = plot_id, value = 1) %>%
    spread(samp_species, value)
 # return(site_samples %>% mutate(study_id = study, plot = plot_id))
}

@@

get_sample_FD <- function(data, n){
  #get random sample for each region and richness level
  plots <- data %>%
    select(site_id, plot) %>%
    distinct()

  sample_occurrence <- pmap_dfr(plots[1:10,], get_plot_sample)

  sample_species_mat <- select(sample_occurrence, -richness, -site_id) %>% #want only species x site info
    select(sort(current_vars())) #sort columns to be in alphabetical order, required for dbFDI()
  sample_trait_mat <- clean_trait_matrix(get_trait_matrix(colnames(sample_species_mat)))

@@

  df <- sample_occurrence %>%
    select(region, richness) %>%
    bind_cols(sample_fd) %>%
    mutate(merge_test = case_when(
      richness == nbsp ~ TRUE,
      TRUE ~ FALSE
    ))

  if(sum(df$merge_test) != dim(df)[1]) warning("richness and region columns may not have been appropriately joined with FD data")

  path <- paste0(here(), "/data/fd_site_samples/", x, "_tmp.tsv.bz2")
  write_tsv(df, path)
  #pb_upload(path)
}
```

```{r}
n <- 100 #sample size

if(!dir.exists(here::here("data", "fd_site_samples"))){
  dir.create(paste0(here::here(), "/data/fd_site_samples"))
  system.time({out <- future_map(1:n, get_sample_FD)})
}

sim <- purrr::map_dfr(fs::dir_ls(path = here::here("data", "fd_samples"), glob="*.tsv.bz2"), readr::read_tsv, .id = "sample")
